---
title: "t2f: Convert Dataframes to LaTeX Tables with Ease"
author: "Ronald G. Thomas"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    highlight: default
vignette: >
  %\VignetteIndexEntry{t2f Package Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, 
  fig.height = 5
)
library(t2f)
```

# Overview

The `t2f` package provides a simple and powerful way to convert R dataframes into professionally formatted LaTeX tables and generate cropped PDF outputs.

## Installation

You can install the package using:

```{r installation, eval=FALSE}
# Install dependencies first
install.packages(c("kableExtra", "glue"))

# Install from GitHub (once published)
# devtools::install_github("your-username/t2f")
```

## Basic Usage

The primary function of the package is `t2f()`, which takes a dataframe and converts it to a LaTeX table with optional styling.

### Simple Example with mtcars Dataset

```{r mtcars-example}
# Use the built-in mtcars dataset
t2f(mtcars, 
    filename = "motor_trend_cars", 
    sub_dir = "tables", 
    scolor = "blue!10", 
    verbose = TRUE)
```

This code will:
- Convert the `mtcars` dataframe to a LaTeX table
- Create a "tables" subdirectory
- Generate a PDF with blue-tinted alternating row colors
- Print progress messages

## Function Parameters

The `t2f()` function offers several parameters:

- `df`: The dataframe you want to convert (required)
- `filename`: Base name for output files (optional, defaults to dataframe name)
- `sub_dir`: Subdirectory for output files (default: "output")
- `scolor`: LaTeX color for row shading (default: "blue!10")
- `verbose`: Print progress messages (default: FALSE)

## Advanced Examples

### Academic Performance Dataset

```{r academic-data-example}
# Create an academic performance dataset
academic_data <- data.frame(
  Student = c("Alice", "Bob", "Charlie"),
  Math = c(95, 88, 92),
  Science = c(90, 85, 93),
  English = c(88, 92, 86)
)

# Convert to LaTeX table with green row shading
t2f(academic_data, 
    filename = "student_scores", 
    scolor = "green!10")
```

### Financial Performance Data

```{r financial-data-example}
# Create a hypothetical financial performance dataset
financial_data <- data.frame(
  Quarter = c("Q1 2023", "Q2 2023", "Q3 2023"),
  Revenue = c(1000000, 1250000, 1500000),
  Expenses = c(800000, 900000, 1000000),
  Profit = c(200000, 350000, 500000)
)

# Convert to LaTeX table with red row shading
t2f(financial_data, 
    filename = "quarterly_financials", 
    scolor = "red!10", 
    verbose = TRUE)
```

## Key Features

### Column Name Sanitization

The package automatically sanitizes column names to ensure compatibility with LaTeX and file systems:

```{r sanitization-example}
# Demonstrate column name sanitization
tricky_names <- data.frame(
  "Column 1!" = 1:3,
  "Special@Column" = 4:6
)

# Automatically converts to valid column names
t2f(tricky_names, filename = "sanitized_columns")
```

### Cell Content Sanitization

Special LaTeX characters like #, %, &, $, are automatically escaped to prevent compilation errors.

## Troubleshooting

### Prerequisites

Ensure the following are installed:
- R
- `kableExtra` package
- `pdflatex`
- `pdfcrop`

### Common Issues
- Check system-wide LaTeX installation
- Verify write permissions in output directory
- Confirm all required LaTeX packages are available

## Conclusion

The `t2f` package simplifies converting R dataframes to professional LaTeX tables, providing:
- Automatic sanitization
- Customizable styling
- Easy PDF generation

## Package Information

- **Version**: 0.1.0
- **Author**: Ronald G. Thomas
- **Dependencies**: kableExtra, glue

## Session Info

```{r session-info}
sessionInfo()
```
